---
title: "HTG miRNA data"
author: "Vrinda Kalia"
date: "4/27/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(openxlsx)
library(edgeR)
```

## Exploring the miRNA data   
This data was generated using matched brain and blood samples obtained from the columbia tumor brain bank.  
MiRNA data was generated from four types of samples:  

*  RNA extracted from brain tissue  

*  Serum  

*  EVs isolated from serum (exoquick)  

*  nEVs isolated from serum (neurodex) 

```{r call in the data, message=FALSE, warning=FALSE}
key <- read.xlsx("/Users/vk2316/Documents/Baccarelli_lab/EVs/Canoll_Bruce/HTG/VLP00932 Sample ID Key.xlsx") %>% 
    mutate(group = as.numeric(factor(Sample.Description)))

raw <- read.xlsx("/Users/vk2316/Documents/Baccarelli_lab/EVs/Canoll_Bruce/HTG/VLP00932_miRNA_PLATE1_20APR2023_Parsed_QualityControlled.xlsx", sheet = "Raw_samples")
IDs <- names(raw[2:24])

raw_t <- raw %>% 
    t() %>% 
    as.data.frame() %>% 
    janitor::row_to_names(1) %>%
    map_df(~as.character(.x)) %>% 
    map_df(~as.numeric(.x)) %>% 
    mutate(ID = IDs) %>% 
    select(ID, everything())
```

## General workflow (Tam 2015)   
1) Short reads are filtered for low-quality sequences/adapters   
2) Remaining sequences are mapped to a reference  
3) Abundance of each biological entity of interest is determined 
4) Resulting count data are normalized  
5) Downstream analysis to probe biological question of interest   

### CPM count-per-million  
```{r normalization CPM, message=FALSE, warning=FALSE}
```

### TMM trimmed mean of m
```{r normalization TMM, message=FALSE, warning=FALSE}
raw_reads <- raw %>% 
    select(-contains("MTC")) %>% 
    filter(!grepl("HK", rna)) %>% 
    filter(!grepl("CTRL", rna)) %>% 
    column_to_rownames(var = "rna")

raw_reads_group <- names(raw_reads) %>% 
    as.data.frame() %>% 
    rename(HTG.EdgeSeq.Run.ID = 1) %>% 
    left_join(., key, by = "HTG.EdgeSeq.Run.ID")

y <- DGEList(counts=raw_reads)
y$samples

# Filter genes with low expression
keep <- filterByExpr(y)
y.drop <- y[keep, , keep.lib.sizes=FALSE]

y.norm <- calcNormFactors(y.drop)
y.norm$samples

y.cpm <- cpm(y.norm, normalized.lib.sizes = TRUE, log = TRUE)

hist(y.cpm[,norm.factor])

y.cpm %>% as.data.frame() %>% 
    rownames_to_column(var = "gene") %>% 
    write_tsv("/Users/vk2316/Documents/Baccarelli_lab/EVs/Canoll_Bruce/HTG/TMM_normalized_log2_cpm.txt")
```




